---
title: "Change Detection Evaluation"
author: "Michael Evans"
date: "`r Sys.Date()`"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    css: custom.css
    social: menu
    df_print: paged
    navbar:
    - {title: "CCI", align: left, href: "https://cci-dev.org"}
    - {title: "Defenders", align: left, href: "http://www.defenders.org"}
    - {title: "", icon: "fa-question-circle fa-lg", align: right, href: "mailto:esa@defenders.org?subject=Change Detection"}
    - {title: "", icon: "fa-github fa-lg", align: right, href: "https://github.com/Defenders-ESC/"}
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("functions.R")

library(dplyr)
library(flexdashboard)
library(MASS)
library(mvtnorm)
library(plotly)
library(pROC)
library(RColorBrewer)
library(rpart)
library(shiny)
library(stringr)
library(tidyr)

IW <- readRDS("data/IW.rds")
IW <- select(IW, systemindex, cv_z, nbr_z, ndsi_z, ndvi_z, ndwi_z, rcvmax_z, habitat, disturbance, Loc, Change)
IW$train <- rbinom(nrow(IW), 1, 0.7)
MAD <- readRDS("data/MAD.rds")
MAD$train <- rbinom(nrow(MAD), 1, 0.7)

IWtree <- rpart(Change ~ cv_z + rcvmax_z + ndvi_z + ndsi_z + ndwi_z + nbr_z,
                data = filter(IW, train ==1), method = 'class')
IWtree_pred <- predict(IWtree, newdata = filter(IW, train == 0))
IWtree_roc <- roc(IW$Change[IW$train ==0], predictor = IWtree_pred[,1])

IWlda <- lda(formula = Change ~ cv_z+nbr_z+ndsi_z+ndvi_z+ndwi_z+rcvmax_z,
               data = filter(IW, train ==1))

#MADlda <- lda(formula = Change ~ chi,
#                  data = MAD)

IWlda_pred <- predict(IWlda, newdata = filter(IW, train == 0))
ldaint <- ldaInt(IWlda, IW[IW$train == 0,], c(2:7))
#MADlda_pred <- predict(MADlda, newdata = MAD)

IWlda_roc <- roc(IW$Change[IW$train ==0], predictor = IWlda_pred$x[,1])

MADlda_roc <- roc(MAD$Change, predictor = MAD$chi)

grid <- expand.grid(habitat = unique(IW$habitat), disturbance = unique(IW$disturbance))

long <- select(IW, Change, habitat, disturbance, cv_z, nbr_z, ndsi_z, ndvi_z, ndwi_z, rcvmax_z)%>%
  mutate(cv_p = 1-pchisq(cv_z, 6),
               rcv_p = 1-pnorm(rcvmax_z),
               ndvi_p = 1-pnorm(ndvi_z),
               ndsi_p = 1-pnorm(ndsi_z),
               ndwi_p = 1-pnorm(ndwi_z),
               nbr_p = 1-pnorm(nbr_z))%>%
  gather("Metric", "Stat", c(4:15))

plot_IW_zs <- function(i){
  a <- grid$habitat[i]
  b <- grid$disturbance[i]
  df <- select(IW, Change, habitat, disturbance, cv_z, rcvmax_z, nbr_z, ndsi_z, ndvi_z, ndwi_z)%>%
    mutate(cv_p = 1-pchisq(cv_z, 6),
               rcv_p = 1-pnorm(rcvmax_z),
               ndvi_p = 1-pnorm(ndvi_z),
               ndsi_p = 1-pnorm(ndsi_z),
               ndwi_p = 1-pnorm(ndwi_z),
               nbr_p = 1-pnorm(nbr_z))%>%
    gather("Metric", "Z", 10:15)%>%
    filter(habitat == a, disturbance == b)
  
  plot_ly(type = "box")%>%
    add_trace(data = df[df$Change == 1,], y = ~Z, x = ~Metric,
              name = "Change", line = list(color = "orange"),
              showlegend = ifelse(i == 1, TRUE, FALSE))%>%
              #legendgroup = ~Change)%>%
    add_trace(data = df[df$Change == 0,], y = ~Z, x = ~Metric,
              name = "No", line = list(color = "blue"),
              showlegend = ifelse(i == 1, TRUE, FALSE))%>%
              #legendgroup = ~Change)%>%
    layout(xaxis = list(title = a, tickangle = 45,
                        tickmode = 'array',
                        tickvals = c('cv_p', 'rcv_p', 'nbr_p', 'ndvi_p',
                                     'ndsi_p', 'ndwi_p'),
                        ticktext = c('CV', 'RCV', 'NBR', 'NDVI', 'NDSI', 'NDWI')),
           yaxis = list(title = b),
           margin = list(b = 100))
}

plot_MAD_zs <- function(i){
  a <- grid$habitat[i]
  b <- grid$disturbance[i]
  df <- filter(MAD, habitat == a, disturbance == b)
  plot_ly(type = "box")%>%
    add_trace(data = df[df$Change == 1,], y = ~chi,
              name = "Change", showlegend = FALSE,
              line = list(color = "orange"))%>%
    add_trace(data = df[df$Change == 0,], y = ~chi,
              name = "No", showlegend = FALSE,
              line = list(color = 'blue'))%>%
    layout(xaxis = list(title = a),
           yaxis = list(title = b),
           margin = list(b = 100))
}

plot_IW_ps <- function(i){
  a <- grid$habitat[i]
  b <- grid$disturbance[i]
  df <- filter(IW, habitat == a, disturbance == b)
  plot_ly(type = "box", data = df, y = ~p, x = ~Change)%>%
    layout(xaxis = list(title = a),
           yaxis = list(title = b))
}

```



Overview
==================================

Column {.sidebar}
-----
### About

```{r readme, echo = FALSE}
fluidPage(
  fluidRow(
    p("The purpose of this app is to evaluate algorithm performance, generically and in specific habitats. Each page will have:"),
    h4("Boxplots"),
    p("Compare the change metric(s) calculated by each algorithm between areas of change and no change.  Ideally, there will be little overlap in values."),
    h4("ROC Curves"),
    p("Help identify threshold values that define change from no change.  Curves plot the power to detect changes (True Positive) against the false positive detection rate at all possible thresholds.  The greater the area under these curves, the better the algorithm is performing."),
    em("Note: the multiple outputs of IW are converted to a single change score using linear discriminant analysis for threshold identification")
  )
)
```

Row
------------------------------------------------------------------------
### Generic change detection using each algorithm.

```{r total_roc, echo = FALSE}
comb <- rbind(gather(MAD, "Metric", "Stat", 9)%>%
                select(Change, Metric, Stat),
              select(long, Change, Metric, Stat)%>%
                filter(grepl("*p", Metric))%>%
                arrange(desc(Change))
                 )

fluidPage(
  fluidRow(
    modalDialog(
    title = h2("Automated Change Detection Algorithm Performance"),
    p("This interactive app presents data produced by Defenders' automated change detection algorithms.  We use two algorithms to identify land cover changes between pairs of satellite images:"),
    tags$ol(
      tags$li("Iteratively re-Weighted land cover change (IW)"),
      tags$li("Multivariate Alteration Detection (MAD)")
      ),
      p("Tabs in the navigation bar at the top of the page let you see data from each algorithm, which produce different outputs."),
    p("Graphs in this app are interactive.  Hover over data to see specific values, and drag to zoom in to parts of a chart."),
    easyClose = TRUE
  ),
    column(6,
           plot_ly(type = "box")%>%
             add_trace(data = comb[comb$Change == 1,], x = ~Metric, y = ~Stat,
                       name = "Change",
                       line = list(color = "orange"))%>%
             add_trace(data = comb[comb$Change == 0,], x = ~Metric, y = ~Stat,
                       name = "No", 
                       line = list(color = "blue"))%>%
             layout(yaxis = list(title = 'P-value'),
                    xaxis = list(title = 'Change metrics',
                                 tickmode = 'array',
                                 tickvals = c('p', 'cv_p', 'rcv_p', 'ndvi_p',
                                              'ndsi_p', 'ndwi_p', 'nbr_p'
                                 ),
                                 ticktext = c('Chi*', 'CV', 'RCV', 'NDVI',
                                              'NDSI', 'NDWI', 'NBR')),
                    title = 'Algorithm outputs change v. no change.',
                    margin = list(b = 100))
           ),

    column(6,

             plot_ly(type = 'scatter', mode = 'lines')%>%
               add_trace(name = "MAD", x = ~1-MADlda_roc$specificities,
                         y = ~MADlda_roc$sensitivities, line = list(color = 'purple'),
                         hoverinfo = 'text',
                         text = ~paste('Chi-square:', MADlda_roc$thresholds,
                             "<br>True Positive:", MADlda_roc$sensitivities,
                             "<br>False Positive:", 1-MADlda_roc$specificities,
                             sep = " "))%>%
               add_trace(name = "IW", x = ~1-IWlda_roc$specificities,
                         y = ~IWlda_roc$sensitivities, line = list(color = 'green'),
                         hoverinfo = 'text',
                         text = ~paste('LDA Score:', IWlda_roc$thresholds,
                                       '<br>True Positive:', IWlda_roc$sensitivities,
                                       '<br>False Positive:',
                                       1-IWlda_roc$specificities))%>%
                   add_annotations(x = 0.5, y = 0.7, xref = "x", yref = "y", showarrow = F,
                      text = paste("CV =", IWlda$scaling[1]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.6, xref = "x", yref = "y", showarrow = F,
                      text = paste("RCV =", IWlda$scaling[6]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.5, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDVI =", IWlda$scaling[4]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.4, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDSI =", IWlda$scaling[3]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.3, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDWI =", IWlda$scaling[5]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.2, xref = "x", yref = "y", showarrow = F,
                      text = paste("NBR =", IWlda$scaling[2]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.1, xref = "x", yref = "y", showarrow = F,
                      text = paste("Int =", ldaint), xanchor = 'left')%>%
               layout(title = "ROC: All Changes",
                      legend = list(x = 0.25, y = 0.2),
                      xaxis = list(title = "False Positive Rate"),
                      yaxis = list(title = "True Positive Rate"))

           )
  )
)
```

Row
------
```{r table, echo = FALSE}
```

IW
===================

Column {.sidebar} 
-----------------
### Examine Subsets

```{r selectors, echo=FALSE}
fluidPage(
  fluidRow(
    p("To display validation data and performance metrics for a specific
      habitat-disturbance type, select from the drop down menus below.
      Then click the 'RUN' button."),
    br(),
    selectInput(inputId = 'habitat', label = "Habitat Type",
                choices = as.list(unique(IW$habitat)),
                     selected = 'forest'),
    br(),
    selectInput(inputId = 'disturbance', label = "Disturbance Type",
                    choices = as.list(unique(IW$disturbance)),
                    selected = 'residential'),
#    br(),
#    selectInput(inputId = 'location', label = "Study Site",
#                choices = as.list(unique(IW$Loc)),
#                selected = 'CityPark'),
    br(),
    actionButton(inputId = 'button', "RUN")
  )
)

  
h <- reactive({input$habitat})
d <- reactive({input$disturbance})

observeEvent(input$habitat, {
  df <- isolate({filter(IW, habitat == h())})
  updateSelectInput(session, "disturbance", label = "Disturbance Type",
                    choices = as.list(unique(df$disturbance)))
})

observeEvent(input$button, {
  #h <- isolate({input$habitat})
  #d <- isolate({input$disturbance})
  iw_sub <- isolate({filter(IW, habitat == h(), disturbance == d())%>%select(Change, habitat, disturbance, cv_z, nbr_z, ndsi_z, ndvi_z, ndwi_z, rcvmax_z, train)})
  # mad_sub <- isolate({filter(MAD, habitat == h(), disturbance == d())})
  df <- isolate({filter(long, habitat == h(), disturbance == d())})
  sig <- isolate({
    cov(filter(IW, Change == 0, habitat == h())[2:6])
  })
  
  mltip <- isolate({
    mutate(IW, p = dmvnorm(IW[,2:6], mean = rep(0,5), sigma = sig))%>%
      filter(habitat == h(), disturbance == d())
  })
  
  output$singlegraph <- renderPlotly({
    plot_ly(type = 'box')%>%
      add_trace(data = df[grepl("*p", df$Metric) & df$Change == 1,], 
            y = ~Stat, x = ~Metric,
            name = "Change", line = list(color = 'orange'))%>%
      add_trace(data = df[grepl("*p", df$Metric) & df$Change == 0,], 
            y = ~Stat, x = ~Metric,
            name = "No", line = list(color= 'blue'))%>%
      layout(title = paste(d(), "in", h()),
             xaxis = list(title = 'Change Metric',
                          tickmode = 'array',
                          tickvals = c('cv_p', 'rcv_p', 'ndvi_p',
                                      'ndsi_p', 'ndwi_p', 'nbr_p'),
                          ticktext = c('CV', 'RCV', 'NDVI', 'NDSI', 'NDWI', 'NBR')),
             yaxis = list(title = 'P-value'),
             legend = list(x = 0.9, y = 1),
             margin = list(b = 100)
             )
  })
  
  # output$MADsingle <- renderPlotly({
  #   plot_ly(type = 'box')%>%
  #     add_trace(data = mad_sub[mad_sub$Change == 1,], y = ~chi,
  #               name = "Change", line = list(color = 'orange'))%>%
  #     add_trace(data = mad_sub[mad_sub$Change == 0,], y = ~chi,
  #               name = "No", line = list(color = 'blue'))%>%    
  #     layout(yaxis = list(title = "Chi-squared"),
  #            legend = list(x = 0.8, y = 1))
  # })
  
  lda_out <- reactive({lda(formula = Change ~ cv_z+nbr_z+ndsi_z+ndvi_z+ndwi_z+rcvmax_z,
               data = filter(iw_sub, train == 1))
  })
  
  lda_pred <- reactive({predict(lda_out(),
                      newdata = filter(iw_sub, train == 0))
  })
  
  lda_roc <- reactive({roc(iw_sub$Change[iw_sub$train == 0], lda_pred()$x[,1])
  })
  
  lda_int <- reactive({ldaInt(lda_out(), iw_sub[iw_sub$train == 0,], 4:9)})
  
  classtree <- reactive({
    rpart(Change ~ cv_z + rcvmax_z + ndvi_z + ndsi_z + ndwi_z + nbr_z,
          data = filter(iw_sub, train == 1), method = 'class')
  })
  
  classtree_pred <- reactive({predict(classtree(), newdata = filter(iw_sub, train == 0))})
  
  classtree_roc <- reactive({roc(iw_sub$Change[iw_sub$train == 0], predictor = classtree_pred()[,1])})
  
  # mad_roc <- reactive({roc(mad_sub$Change, mad_sub$chi)})
  
#  output$lda <- renderPlotly({
#    df <- data.frame(Change = iw_sub$Change, Pred = lda_pred()$x[,1])
#    plot_ly(type = 'histogram')%>%
#      add_trace(data = df[df$Change == 1, ],
#            x = ~Pred, name = "Change", histnorm = 'probability',
#            marker = list(color = 'orange'), opacity = 0.6)%>%
#      add_trace(data = df[df$Change == 0, ],
#                x = ~Pred, name = "No", histnorm = 'probability',
#                marker = list(color = 'blue'), opacity = 0.6)%>%
#      layout(barmode = 'overlay',
#             legend = list(x = 0.75, y = 1),
#             xaxis = list(title = "LDA Score"))
#  })
  
  output$lda <- renderPlotly({
    df <- data.frame(Change = iw_sub$Change[iw_sub$train == 0], Pred = lda_pred()$x[,1])
    plot_ly(type = 'box', orientation = 'h')%>%
      add_trace(data = df[df$Change == 1, ], x = ~Pred,
            y = ~Change, name = "Change",
            line = list(color = 'orange'),
            boxpoints = "all", pointpos = 0)%>%
      add_trace(data = df[df$Change == 0, ], x = ~Pred,
                y = ~Change, name = "No",
                line = list(color = 'blue'),
                boxpoints = 'all', pointpos = 0)%>%
      layout(title = paste("LDA:", d(), "x", h()),
             legend = list(x = 0.2, y = 1),
             xaxis = list(title = "LDA Score"))
  })
  
#  output$madlda <- renderPlotly({
#    plot_ly(type = 'histogram')%>%
#      add_trace(data = mad_sub[mad_sub$Change == 1, ],
#            x = ~chi, name = "Change", histnorm = 'probability',
#            marker = list(color = 'orange'), opacity = 0.6)%>%
#      add_trace(data = mad_sub[mad_sub$Change == 0, ],
#                x = ~chi, name = "No", histnorm = 'probability',
#                marker = list(color = 'blue'), opacity = 0.6)%>%
#      layout(barmode = 'overlay',
#             legend = list(x = 0.75, y = 1),
#             xaxis = list(title = "Chi-squared"))
#  })
  
  # output$madlda <- renderPlotly({
  #   plot_ly(type = 'box', orientation = 'h')%>%
  #     add_trace(data = mad_sub[mad_sub$Change == 1, ], x = ~chi,
  #           y = ~Change, name = "Change",
  #           line = list(color = 'orange'),
  #           boxpoints = "all", pointpos = 0)%>%
  #     add_trace(data = mad_sub[mad_sub$Change == 0, ], x = ~chi,
  #               y = ~Change, name = "No",
  #               line = list(color = 'blue'),
  #               boxpoints = 'all', pointpos = 0)%>%
  #     layout(
  #            legend = list(x = 0.75, y = 1),
  #            xaxis = list(title = "Chi-squared"))
  # })
  
#  output$madldaroc <- renderPlotly({
#    plot_ly()%>%
#      add_trace(x = ~1-mad_roc()$specificities, y = ~mad_roc()$sensitivities,
#                type = 'scatter', mode = 'lines',
#                hoverinfo = 'text',
#                text = ~paste('Chi-square:', mad_roc()$thresholds,
#                             "<br>True Positive:", mad_roc()$sensitivities,
#                             "<br>False Positive:", 1-mad_roc()$specificities,
#                             sep = " "))%>%
    #   layout(title = paste("ROC:", d(), "x", h(), sep = " "),
    #          xaxis = list(title = "False Positive Rate"),
    #                       #autorange = 'reversed'),
    #          yaxis = list(title = "True Positive Rate"))
    # })
  
  
  output$ldaroc <- renderPlotly({
    plot_ly()%>%
      add_trace(x = ~1-lda_roc()$specificities, y = ~lda_roc()$sensitivities,
                type = 'scatter', mode = 'lines',
                hoverinfo = 'text',
                text = ~paste('LDA Score:', lda_roc()$thresholds,
                             "<br>True Positive:", lda_roc()$sensitivities,
                             "<br>False Positive:", 1-lda_roc()$specificities,
                             sep = " "))%>%
      add_annotations(x = 0.5, y = 0.7, xref = "x", yref = "y", showarrow = F,
                      text = paste("CV =", lda_out()$scaling[1]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.6, xref = "x", yref = "y", showarrow = F,
                      text = paste("RCV =", lda_out()$scaling[6]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.5, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDVI =", lda_out()$scaling[4]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.4, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDSI =", lda_out()$scaling[3]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.3, xref = "x", yref = "y", showarrow = F,
                      text = paste("NDWI =", lda_out()$scaling[5]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.2, xref = "x", yref = "y", showarrow = F,
                      text = paste("NBR =", lda_out()$scaling[2]), xanchor = 'left')%>%
            add_annotations(x = 0.5, y = 0.1, xref = "x", yref = "y", showarrow = F,
                      text = paste("Int =", lda_int()), xanchor = 'left')%>%
      
      layout(title = paste("LDA ROC:", d(), "x", h(), sep = " "),
             xaxis = list(title = "False Positive Rate"),
             yaxis = list(title = "True Positive Rate"))
    })
  
  bayes_roc <- roc(mltip$Change, mltip$p)
  
  output$bayes <- renderPlotly({
    plot_ly(type = 'box', orientation = 'h')%>%
      add_trace(data = mltip[mltip$Change == 1, ], x = ~p, y = ~Change,
                line = list(color = 'orange'), name = 'Change',
                boxpoints = 'all', pointpos = 0)%>%
      add_trace(data = mltip[mltip$Change == 0, ], x = ~p, y = ~Change,
                line = list(color = 'blue'), name = 'No',
                boxpoints = 'all', pointpos = 0)%>%
      layout(#barmode = 'overlay',
        title = paste("Bayesian:", d(), 'x', h()),
        legend = list(x = 0.75, y = 1),
        xaxis = list(title = "P-value"),
        margin = list(b = 100))
  })
                 
  output$bayesroc <- renderPlotly({
    plot_ly()%>%
      add_trace(x = ~1-bayes_roc$specificities, y = ~bayes_roc$sensitivities,
                type = 'scatter', mode = 'lines',
                hoverinfo = 'text',
                text = ~paste('P-value:', bayes_roc$thresholds,
                              '<br>True Positive:', bayes_roc$sensitivities,
                              '<br>False Positive:', 1-bayes_roc$specificities,
                              sep = " "))%>%
      layout(title = paste("Bayesian ROC:", d(), "x", h(), sep = " "),
             xaxis = list(title = "False Positive Rate"),
             yaxis = list(title = "True Positive Rate"),
             margin = list(b = 100))
    })

  output$classt <- renderPlotly({
    df <- data.frame(Change = iw_sub$Change[iw_sub$train == 0], Pred = classtree_pred()[,1])
    plot_ly(type = 'box', orientation = 'h')%>%
      add_trace(data = df[df$Change == 1, ], x = ~Pred,
            y = ~Change, name = "Change",
            line = list(color = 'orange'),
            boxpoints = "all", pointpos = 0)%>%
      add_trace(data = df[df$Change == 0, ], x = ~Pred,
                y = ~Change, name = "No",
                line = list(color = 'blue'),
                boxpoints = 'all', pointpos = 0)%>%
      layout(title = paste("Classification:", d(), "x", h()),
             legend = list(x = 0.2, y = 1),
             xaxis = list(title = "Classification Score"))
  })
  
    output$classroc <- renderPlotly({
    plot_ly()%>%
      add_trace(x = ~1-classtree_roc()$specificities, y = ~classtree_roc()$sensitivities,
                type = 'scatter', mode = 'lines',
                hoverinfo = 'text',
                text = ~paste('Class Threshold:', classtree_roc()$thresholds,
                             "<br>True Positive:", classtree_roc()$sensitivities,
                             "<br>False Positive:", 1-classtree_roc()$specificities,
                             sep = " "))%>%
      layout(title = paste("Classification Tree ROC:", d(), "x", h(), sep = " "),
             xaxis = list(title = "False Positive Rate"),
             yaxis = list(title = "True Positive Rate"))
    })
})

```

Row 
--------------------------------------------------------------------------
### IW Algorithm Output: 5 metrics from areas of change and no change.

```{r iw_all, echo = FALSE}

fluidPage(
  fluidRow(
    column(6,
           renderPlotly({
             subplot(lapply(1:nrow(grid), plot_IW_zs),
                   nrows = length(unique(IW$disturbance)), shareY = TRUE, titleY = TRUE,
                   shareX =TRUE,titleX= TRUE)%>%
             layout(title = "Habitat by Disturbance Combos")
           })
    ),
    
    column(6,
           plotlyOutput('singlegraph')
           )
  )
)
```

Row 
--------------------------------------------------------------------------
### Change/No-Change Thresholds: Linear Discriminant Analysis (LDA)

LDA transforms multiple measures into a single score used to discriminate among groups.  The ROC curve compares true positive and false positive detection rates for different LDA thresholds.

```{r iw_lda, echo=FALSE}

#fluidPage(
  fluidRow(
    column(5,
           plotlyOutput("lda")
           ),
    column(1),
    column(5,
           plotlyOutput("ldaroc")
           )
  )
#)

```

Row
----
### Change/No-Change Thresholds: Bayesian Classification

Bayesian classification is another approach used to convert multivariate output to a single measure (p-value) to discriminate between change and no change.
```{r bayesian, echo = FALSE}

fluidRow(
  column(5,
         plotlyOutput('classt')
         ),
  column(1),
  column(5,
         plotlyOutput('classroc')
         )
)
```

MAD
==================================

Column {.sidebar} 
-----------------
### Examine Subsets

```{r mad_selectors, echo=FALSE}
fluidPage(
  fluidRow(
    p("To display validation data and performance metrics for a specific
      habitat-disturbance type, select from the drop down menus below.
      Then click the 'RUN' button."),
    br(),
    selectInput(inputId = 'mad_habitat', label = "Habitat Type",
                choices = as.list(unique(IW$habitat)),
                     selected = 'forest'),
    br(),
    selectInput(inputId = 'mad_disturbance', label = "Disturbance Type",
                    choices = as.list(unique(IW$disturbance)),
                    selected = 'residential'),
#    br(),
#    selectInput(inputId = 'location', label = "Study Site",
#                choices = as.list(unique(IW$Loc)),
#                selected = 'CityPark'),
    br(),
    actionButton(inputId = 'mad_button', "RUN")
  )
)

  
mh <- reactive({input$mad_habitat})
md <- reactive({input$mad_disturbance})

observeEvent(input$mad_habitat, {
  df <- isolate({filter(MAD, habitat == mh())})
  updateSelectInput(session, "mad_disturbance", label = "Disturbance Type",
                    choices = as.list(unique(df$disturbance)))
})

observeEvent(input$mad_button, {
  #h <- isolate({input$habitat})
  #d <- isolate({input$disturbance})
  #iw_sub <- isolate({filter(IW, habitat == h(), disturbance == d())})
  mad_sub <- isolate({filter(MAD, habitat == mh(), disturbance == md())})
  #df <- isolate({filter(long, habitat == h(), disturbance == d())})
  #sig <- isolate({
  #  cov(filter(IW, Change == 0, habitat == h())[2:6])
  #})
  
  #mltip <- isolate({
  #  mutate(IW, p = dmvnorm(IW[,2:6], mean = rep(0,5), sigma = sig))%>%
  #    filter(habitat == h(), disturbance == d())
  #})
  
  #output$singlegraph <- renderPlotly({
  #  plot_ly(type = 'box')%>%
  #    add_trace(data = df[grepl("*p", df$Metric) & df$Change == 1,], 
  #          y = ~Stat, x = ~Metric,
  #          name = "Change", line = list(color = 'orange'))%>%
  #    add_trace(data = df[grepl("*p", df$Metric) & df$Change == 0,], 
  #          y = ~Stat, x = ~Metric,
  #          name = "No", line = list(color= 'blue'))%>%
  #    layout(title = paste(d(), "in", h()),
  #           xaxis = list(title = 'Change'),
  #           yaxis = list(title = 'P-value'),
  #           legend = list(x = 0.9, y = 1)
  #           )
  #})
  
  output$MADsingle <- renderPlotly({
    plot_ly(type = 'box')%>%
      add_trace(data = mad_sub[mad_sub$Change == 1,], y = ~chi,
                name = "Change", line = list(color = 'orange'))%>%
      add_trace(data = mad_sub[mad_sub$Change == 0,], y = ~chi,
                name = "No", line = list(color = 'blue'))%>%    
      layout(yaxis = list(title = "Chi-squared"),
             legend = list(x = 0.8, y = 1),
             title = paste(md(), "in", mh()))
  })
  
  #lda_out <- reactive({lda(formula = Change ~ rcvmax_z+cv_z+nbr_z+ndsi_z+ndvi_z,
  #             data = iw_sub)
  #})
  
  #lda_pred <- reactive({predict(lda_out(),
  #                    newdata = iw_sub)
  #})
  
  #lda_roc <- reactive({roc(iw_sub$Change, lda_pred()$x[,1])
  #})
  
  mad_roc <- reactive({roc(mad_sub$Change, mad_sub$chi)})
  
#  output$lda <- renderPlotly({
#    df <- data.frame(Change = iw_sub$Change, Pred = lda_pred()$x[,1])
#    plot_ly(type = 'histogram')%>%
#      add_trace(data = df[df$Change == 1, ],
#            x = ~Pred, name = "Change", histnorm = 'probability',
#            marker = list(color = 'orange'), opacity = 0.6)%>%
#      add_trace(data = df[df$Change == 0, ],
#                x = ~Pred, name = "No", histnorm = 'probability',
#                marker = list(color = 'blue'), opacity = 0.6)%>%
#      layout(barmode = 'overlay',
#             legend = list(x = 0.75, y = 1),
#             xaxis = list(title = "LDA Score"))
#  })
  
  #output$lda <- renderPlotly({
  #  df <- data.frame(Change = iw_sub$Change, Pred = lda_pred()$x[,1])
  #  plot_ly(type = 'box', orientation = 'h')%>%
  #    add_trace(data = df[df$Change == 1, ], x = ~Pred,
  #          y = ~Change, name = "Change",
  #          line = list(color = 'orange'),
  #          boxpoints = "all", pointpos = 0)%>%
  #    add_trace(data = df[df$Change == 0, ], x = ~Pred,
  #              y = ~Change, name = "No",
  #              line = list(color = 'blue'),
  #              boxpoints = 'all', pointpos = 0)%>%
  #    layout(
  #           legend = list(x = 0.75, y = 1),
  #           xaxis = list(title = "LDA Score"))
  #})
  
#  output$madlda <- renderPlotly({
#    plot_ly(type = 'histogram')%>%
#      add_trace(data = mad_sub[mad_sub$Change == 1, ],
#            x = ~chi, name = "Change", histnorm = 'probability',
#            marker = list(color = 'orange'), opacity = 0.6)%>%
#      add_trace(data = mad_sub[mad_sub$Change == 0, ],
#                x = ~chi, name = "No", histnorm = 'probability',
#                marker = list(color = 'blue'), opacity = 0.6)%>%
#      layout(barmode = 'overlay',
#             legend = list(x = 0.75, y = 1),
#             xaxis = list(title = "Chi-squared"))
#  })
  
  output$madlda <- renderPlotly({
    plot_ly(type = 'box', orientation = 'h')%>%
      add_trace(data = mad_sub[mad_sub$Change == 1, ], x = ~chi,
            y = ~Change, name = "Change",
            line = list(color = 'orange'),
            boxpoints = "all", pointpos = 0)%>%
      add_trace(data = mad_sub[mad_sub$Change == 0, ], x = ~chi,
                y = ~Change, name = "No",
                line = list(color = 'blue'),
                boxpoints = 'all', pointpos = 0)%>%
      layout(
             legend = list(x = 0.75, y = 1),
             xaxis = list(title = "Chi-squared"))
  })
  
  output$madldaroc <- renderPlotly({
    plot_ly()%>%
      add_trace(x = ~1-mad_roc()$specificities, y = ~mad_roc()$sensitivities,
                type = 'scatter', mode = 'lines',
                hoverinfo = 'text',
                text = ~paste('Chi-square:', mad_roc()$thresholds,
                             "<br>True Positive:", mad_roc()$sensitivities,
                             "<br>False Positive:", 1-mad_roc()$specificities,
                             sep = " "))%>%
      layout(title = paste("ROC:", md(), "x", mh(), sep = " "),
             xaxis = list(title = "False Positive Rate"),
                          #autorange = 'reversed'),
             yaxis = list(title = "True Positive Rate"))
    })
  
  
  #output$ldaroc <- renderPlotly({
  #  plot_ly()%>%
  #    add_trace(x = ~1-lda_roc()$specificities, y = ~lda_roc()$sensitivities,
  #              type = 'scatter', mode = 'lines',
  #              hoverinfo = 'text',
  #              text = ~paste('LDA Score:', lda_roc()$thresholds,
  #                           "<br>True Positive:", lda_roc()$sensitivities,
  #                           "<br>False Positive:", 1-lda_roc()$specificities,
  #                           sep = " "))%>%
  #    layout(title = paste("ROC:", d(), "x", h(), sep = " "),
  #           xaxis = list(title = "False Positive Rate"),
  #           yaxis = list(title = "True Positive Rate"))
  #  })
  
  #bayes_roc <- roc(mltip$Change, mltip$p)
  
  #output$bayes <- renderPlotly({
  #  plot_ly(type = 'box', orientation = 'h')%>%
  #    add_trace(data = mltip[mltip$Change == 1, ], x = ~p, y = ~Change,
  #              line = list(color = 'orange'), name = 'Change',
  #              boxpoints = 'all', pointpos = 0)%>%
  #    add_trace(data = mltip[mltip$Change == 0, ], x = ~p, y = ~Change,
  #              line = list(color = 'blue'), name = 'No',
  #              boxpoints = 'all', pointpos = 0)%>%
  #    layout(barmode = 'overlay',
  #           legend = list(x = 0.75, y = 1),
  #           xaxis = list(title = "P-value"))
  #})
                 
  output$bayesroc <- renderPlotly({
    plot_ly()%>%
      add_trace(x = ~1-bayes_roc$specificities, y = ~bayes_roc$sensitivities,
                type = 'scatter', mode = 'lines',
                hoverinfo = 'text',
                text = ~paste('P-value:', bayes_roc$thresholds,
                              '<br>True Positive:', bayes_roc$sensitivities,
                              '<br>False Positive:', 1-bayes_roc$specificities,
                              sep = " "))%>%
      layout(title = paste("Bayesian:", md(), "x", mh(), sep = " "),
             xaxis = list(title = "False Positive Rate"),
             yaxis = list(title = "True Positive Rate"))
    })
})


```

Row 
--------------------------------------------------------------------------
### MAD Algorithm Output: Chi-square statistic in areas of change and no change.

```{r mad_all, echo = FALSE}
fluidPage(
  fluidRow(
    column(6,
           renderPlotly({
             subplot(lapply(1:nrow(grid), plot_MAD_zs),
                   nrows = length(unique(MAD$disturbance)), shareY = TRUE, titleY = TRUE,
                   shareX =TRUE,titleX= TRUE)%>%
             layout(title = "Habitat x Disturbance Combos")
           })
    ),
    
    column(6,
           plotlyOutput('MADsingle')
           )
  )
)
```

Row
--------------------------------------------------------------------------
### Change/No-Change Thresholds

The MAD algorithm produces a single measure (Chi-squared statistic) that we can use to discriminate between change and no change.

```{r mad_lda, echo=FALSE}

fluidPage(
  fluidRow(
    column(5,
           plotlyOutput("madlda")
           ),
    column(1),
    column(5,
           plotlyOutput("madldaroc")
           )
  )
)

```

